### 2024-04-17 presentations/understanding the problem.qmd
commutativity

---
title: "Understanding the problem"
format: 
  revealjs:
    smaller: true
    scrollable: true
editor: visual
---

```{r setup}
library(knitr)
```

## Blåbærkrisa

::: columns
::: {.column width="30%"}
Same data gave different indicator values
:::

::: {.column width="70%"}
```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example1.R")
example1
```

::: aside
Indicators with grey background
:::
:::
:::

## This project

-   250 000 kr

-   166 hours

-   1 month

-   Manuscript on pre-print server

-   Produce guidelines

------------------------------------------------------------------------

Example of indeces at EA scale aggregated to EAA scale

```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example3.R")
example3
```

------------------------------------------------------------------------

Example of indicators aggregated at EEA scale, then turned into an index

```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example4.R")
example4
```

------------------------------------------------------------------------

Adding area weighting to the first example

```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example3b.R")
example3b
```

------------------------------------------------------------------------

and to the second example

```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example5.R")
example5
```

## Commutativty

These are examples of cummutative aggregation workflows

![](img/commutativity.PNG)

------------------------------------------------------------------------

Next, variables aggregated at EAA scale by taking the area weighted mean. When would this make sense to do? Obviously not for count variables.

```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example6.R")
example6
```

------------------------------------------------------------------------

Here, variable aggregated at EAA scale with no area weighting

```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example7.R")
example7
```

------------------------------------------------------------------------

Example with two indicators. Scaling indicators at the sub-regions also makes it possible to calculate an index for the sub-regions. But to get an index for the whole region, should we aggregate the

```{r}
#| fig-align: "left"
#| fig-height: 6
#| fig-width: 7
source("../R/example2.R")
example2
```