### 2025-09-17 manuscript/table_bilberry.qmd

```{r}
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r dat}
#| echo: false
dat <- tibble(
    " " = c(
        "Variable values",
        "$X_{100}$",
        "Indicator values"),
    "Plot 1" = c(
        "2",
        "20*",
        "0.1*"),
    "Plot 2" = c(
        "36",
        "20*",
        "1.0*"),
    "EA" = c(
        "19",
        "20",
        "0.95 (0.55*)"))
```

```{r tbl-bilberry}
#| echo: false
#| tbl-cap: "Dummy data for the bilberry example. Variable values (% cover of bilberry) are recorded for each vegetation plot, and aggregated (arithmetic mean) to the ecosystem asset (EA) level. The X~100~ value is defined for the EA level, but in this hypothetical case it is also dis-aggregated to each vegetation plot and used for normalising variable values also at the plot level. The dis-aggregated X~100~ values, and subsequent values that rely on these, are marked with an asterisk (*)."

kbl(dat,
    format = "latex",
    booktabs = TRUE,
    escape = F
    ) 
```